Making multiple-locus association mapping on a genome-wide scale routine

### 2017-11-09 AM+draft.tex
09 Nov 2017  lunch time commit

>>>>

Of the multiple-locus applications considered in this paper, 
MLMM \cite{segura2012efficient} bares the greatest similarity to Eagle algorithmically.  MLMM, like Eagle, performs 
association mapping by building the "best" linear mixed model.  The model used for analysis 
is the same as (\ref{eq1}) except that the genetic effect is 
\begin{equation}
\label{eqmlmm}
u_g = \sum_{k=1}^s  m_{S_k} a_{S_k} + g
\end{equation}
where 
$g^{(n ; \times 1)}$ is a vector of random polygenic effects with distribution 
$g \sim N(0, \sigma^2_g K)$. The model is built iteratively, via forward-backward variable selection.  In Eagle, at each iteration, 
the snp in strongest association with the trait is found by calculating the score statistic $t_j^2$.  In MLMM, at each iteration, 
a separate linear mixed model analysis is performed for each snp not already selected. The snp yielding the most 
significant result (lowest p-value) is selected for inclusion in the fixed effects part of (\ref{eqmlmm}).  
Two selection 
criteria are available for the model building process: 
the EBIC and the multiple-Bonferroni criterion. We used only the EBIC.

\textbf{r2VIM:} R2VIM is an R package specifically designed for association mapping \cite{szymczak2016r2vim}. 
It implements random forests and a new way of measuring the importance of a snp.  In random forests, the worth of a predictor is measured, empirically, by calculating its importance score. It is from these importance scores that snp can be ordered in terms of their strength of association with a trait. The challenge though is in knowing what proportion of the 
highest ordered snp are in true association. R2VIM addresses this by calculating a relative importance score for a snp. Briefly,
$m$ random forest analyses of the data are performed. Each analysis is reliant upon a different random number seed. 
From an analysis, a relative importance score for a snp is obtained by taking the ratio of its importance score to the absolute 
minimum of the importance scores across all the snp.  Those 
snp with high relative scores across all $m$ analyses are deemed to be of interest \cite{szymczak2016r2vim}.
Applying r2VIM to the analysis of GWAS data is not without its challenges.  First, it is not possible to force every 
decision tree to have important (fixed) effects without a change to the internal workings of the r2VIM package. 
So we adopted a two step 
strategy. In the first step, the fixed effects are regressed on to the trait and the residuals recorded.  In the second step, 
r2VIM is employed were the trait data are the residuals obtained from the first step. This is not ideal (ref to problems with two
step approaches).  Second, there is no relationship between relative importance scores and significance levels. Consequently, 
setting an appropriate genome-wide threshold is problematic. We avoid this problem by focusing, in the simulation study, 
only on the relationship between threshold level and its false discovery rate and power.
